\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-02-26 papers/methods/paper.tex
Appendix update

We assume that our selection effects can be reasonably well encapsulated by independent functions of (actual) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$. Our denominator then becomes
\begin{align}
w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta),
\end{align}
where for simplicity we have not written out all the integrals which do not interact with the selection effects explicitly. Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$. Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{align}
w = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{align}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernovae redshifts, we can approximate the integral over redshift by evaluating
\begin{align}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B
\end{align}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{align}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{align}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.

@@

Considering a $3\times3$ matrix with cholesky decomposition
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
\end{pmatrix},
\end{align}
the original covariance matrix $\Sigma$ is given by
\begin{align}
\Sigma = \begin{pmatrix}
a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
\end{pmatrix}.
\end{align}
Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered colour parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$. We can now see that it is possible to add extra uncertainty to the colour observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix. To take our covariance and add on the diagonal uncertainty for colour an extra $\sigma_e$ term, we get
\begin{align}
C = \begin{pmatrix}
\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
\end{pmatrix}
\end{align}.
The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
\end{pmatrix},
\end{align}
where $g = \sqrt{f^2 + \sigma_e^2} - f$. This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent colour uncertainty. For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.

>>>>

We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$. Our denominator then becomes
\begin{align}
w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta), \label{eq:sel}
\end{align}
where for simplicity we have not written out all the integrals which do not interact with the selection effects explicitly. Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$ (which is why they were not included in equation \eqref{eq:sel}). Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{align}
w = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{align}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernovae redshifts, we can approximate the integral over redshift by evaluating
\begin{align}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{align}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{align}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
\end{align}
Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.

@@

Considering a $3\times3$ matrix with cholesky decomposition
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
\end{pmatrix},
\end{align}
the original covariance matrix $\Sigma$ is given by
\begin{align}
\Sigma = \begin{pmatrix}
a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
\end{pmatrix}.
\end{align}
Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered colour parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$. We can now see that it is possible to add extra uncertainty to the colour observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix. To take our covariance and add on the diagonal uncertainty for colour an extra $\sigma_e$ term, we get
\begin{align}
C = \begin{pmatrix}
\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
\end{pmatrix}.
\end{align}
The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
\begin{align}
L = \begin{pmatrix}
a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
\end{pmatrix},
\end{align}
where $g = \sqrt{f^2 + \sigma_e^2} - f$. This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent colour uncertainty. For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.
